Ribosome accumulation during early phase resistance training

### 2020-08-24 manuscript.Rmd
Small changes

## Study overview

## Muscle strength and thickness
Muscle strength was assessed as maximal knee extension isokinetic (90&deg; sec<sup>-1</sup>) and isometric (60&deg; angle, fully extended leg 0&deg;) torque. After a brief warm-up (5-min cycling, RPE 12-14), participants were seated and secured in the individually adjusted dynamometer. The test protocol consisted of unilateral knee-extensions. Participants were instructed to gradually increase their effort during three warm-up repetitions (50, 60 and, 70% of subjective maximal effort). After a 30-sec restitution period participants were instructed to perform three repetitions with maximal effort in the concentric phase. Sixty seconds after the isokinetic test the lever automatically moved to a 60&deg; angle and participants were instructed to push against the lever enough to see feedback from the visual feedback system. After an additional 15-sec restitution period, participants were instructed to push against the lever with maximal effort. During the same assessment session, participants remained seated in the dynamometer for meassurement from both legs. The first measurement was alternated between legs and sessions. For statistical treatment of the data, all successful attempts were used. Successful attempts were decided based on participants (...). The last strength assessment at baseline was performed at least **seven** days prior to the first biopsy sampling. At least one strength was performed on separate days with two sessions allowed perform on the same day with a short rest between sessions. Post training assessments were performed approximately 48 hours after the last session and after seven days of no training.

Muscle thickness was assessed using (...) ultra sound.

>>>>

## Study overview
Eighteen volunteers were recruited to the study. Eligible participants had to be non-smokers, between 18 and 35 years of age, report a training history of less than one RT session per week during the six months leading up to the study and not consume dietary supplements or medication with known effects on muscle metabolism. Potential participants were further screened for injuries causing impaired strength and/or affecting their ability to perform RT, symptoms or history of disease or known adverse reactions to local anaesthetics. Included participants were allocated to either an experimental group (EXP, *n* = 11) or a non-training control group (CTRL, *n* = 8; see Table 1 for participant characteristics). EXP performed a 12 session RT protocol consisting unilateral knee-extension to allow for within-participant comparison exercise volume. In EXP, one leg performed constant volume (CONST, 6 sets per session) and the other leg performed variable volume over the course of the study (VAR, four sessions each with 6, 3 and 9 sets per session; Figure 1A). CTRL did not partake in RT and were instructed to continue their everyday activities. Muscle biopsies were sampled bilaterally in EXP before and 48 h after the first session and 48 h after the fourth, fifth, eight, ninth and twelfth session as wells as after a seven day de-training period. Muscle biopsies were obtained from CTRL at three occasions with 48 h separating the first two and 2-4 weeks separating second and third sampling. EXP and CTRL performed strength assessments at least seven days prior to the first biopsy sampling and 72 h after the twelwth session (EXP) and 24 h after the last biopsy (EXP and CTRL). Appendicular lean mass (Dual-energy X-ray absorptiometry, DXA) and muscle thickness of *m. vastus lateralis* were assessed prior to the first biopsy (EXP and CTRL) as wells as before the second to last (EXP) and last (EXP and CTRL) biopsy (Figure 1A).

## RT protocol
Prior to all RT-sessions, participants performed a standardized warm-up consisting of 5 min ergometer cycling (RPE 12-14), followed by ten repetitions each of push-ups, sit-ups and back-extensions. After warm-up participants performed unilateral knee-extension with the prescribed number of sets. Each set was prescribed with 10 repetitions maximum (RM). When sets was completed with either fewer (8) or more (12) repetitions than 10, the resistance was adjusted accordingly. Inter-set rest periods were 90 sec. Participants performed the first set alternating between the right and left leg every other session and the set of the contralateral leg was performed in the rest-period of the first leg. The second session of each four-session block (session 2, 6 and 10) were performed at a sub-maximal resistance (~90% of the previous session) with the same number of repetitions. In each session, after the knee-extension exercise, participants completed two sets each of three upper-body exercises (bench press, lateral pull-down and shoulder press). Prior to each session, participants were asked to rate muscle soreness on a scale from 0 to 10 (0 = no soreness, 10 = hurting very bad, e.g. not being able to walk properly) [@Vickers2010]. After each session, participants were asked to rate their perceived exertion (Borg CR10). After the completion of each session participants were given a standardized drink to aid recovery (0.15 g × kg-1 protein, 11.2 g × kg-1 carbohydrates and 0.5 g × kg-1 fat).

## Muscle strength, body composition and muscle thickness assessments
Muscle strength was assessed as maximal voluntary isokinetic (90&deg; sec<sup>-1</sup>) and isometric (60&deg; angle, fully extended leg 0&deg;) knee extension torque. After a brief warm-up (5-min cycling, RPE 12-14), participants were seated and secured in the individually adjusted dynamometer. Participants were instructed to gradually increase their effort during three warm-up repetitions (50, 60 and, 70% of subjective maximal effort). After a 30-sec restitution period participants were instructed to perform three repetitions with maximal effort in the concentric phase. Sixty seconds after the isokinetic test the lever automatically moved to a 60&deg; angle and participants were instructed to push against the lever enough to see feedback from the visual feedback system. After an additional 15-sec restitution period, participants were instructed to push against the lever with maximal effort. During the same assessment session, participants remained seated in the dynamometer for meassurement from both legs. The first measurement was alternated between legs and sessions. For statistical treatment of the data, all successful attempts were used. Successful attempts were decided based on participants (...). The last strength assessment at baseline was performed at least **seven** days prior to the first biopsy sampling. At least one strength was performed on separate days with two sessions allowed perform on the same day with a short rest between sessions. Post training assessments were performed approximately 48 hours after the last session and after seven days of no training.

Muscle thickness was assessed using (...) ultra sound. DXA.
